On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-04-05 TTau_Halpha/revision01/response_to_ref.txt

      	Since the stream lines in our simple disc wind are represented
	by straight lines, the wind never bends toward the symmetry
	axis (z-axis) as the distance from the disc increases unlike
	the model of Blandford & Payne (1982). We therefore used a
	larger launching angle to simulate the stream lines in the far
	field (at a larger distance from the disc plane) in their
	model.  This is an intrinsic problem with approximating the
	stream lines with straight lines. A better model is perhaps to
	compute the field lines using Blandford & Payne (1982) model
	as done so in Alencar et al (2005, A&A, 440, 595).  Also, in
	our mode, if the launching angle is too small (<60 deg.), the
	wind absorption feature would be not been seen or would be
	very small for profiles computed with i<60 deg.  If the
	launching angle is 60 deg. at the inner most part of the wind,
	and the inclination angle (measured from the symmetry axis,
	z-axis) is less than 60 deg., little (wind) absorption is
	expected, as demonstrated in Figure 7.  The situation will
	however changes if the field/stream line is not straight and
	bending towards the symmetry axis as the wind travel higher
	above the disc.  Since more than 50 per cent of observed
	H-alpha profiles (e.g. Reipurth 1996, A&AS, 120, 229 and
	Alencar & Basri 2000, ApJ 119, 1881) are affected by wind
	absorption, it was necessary for our model to used the
	launching angle (at the inner most part) greater than 60
	degrees to be more consistent with observations; although this
	would mean that the launching angle condition of Blandford and
	Payne (1982) is violated for wind from a very inner part of
	the disc.

	 In addition to the models given in Figs. 6 and 9, we have
	 computed the models with d=5Rstar (not shown here) which
	 satisfies the wind launching condition of Blandford & Payne
	 (1982) in all parts of the disc wind zone in our model.
	 These models showed that the wind absorption and emission
	 features weaken, but overall morphology dependence of the
	 profiles on the wind temperature rate and the wind
	 acceleration rate is similar to what we found in the models
	 shown in Figs. 6 and 9.  The minimum inclination angle at
	 which the wind absorption becomes important is larger than
	 that of the models with a larger source displacement (d = 22
	 Rstar) as expected."

Sect. 4.3, 1st paragraph: You say the parameters used for the
magnetosphere are the same as in Fig.5, but in that figure you have
models with 3 different inclinations. Did you calculate models for the
three inclinations or just for i=55 degrees ? Please specify this in
the text.

   * All the models in Section 4 are computed at i=55 deg. The
     profiles computed at 10 and 80 deg. are given in Appendix A. 
     I see that is confusing. 
     I corrected the sentence as follows:

     ORIGINAL
        "The parameters used for the magnetosphere are  same as in 
         Fig. 5, ...."

     REVISED
        "The parameters used for the magnetosphere are same as in 
         Fig. 5 (with i=55 deg.), ...."

   * Also the last sentence is the same paragraph changed to reflect
     your comment:

     ORIGINAL
       "Fig. 9 shows the models profiles computed 
        using the same ranges of the disc-wind temperature and wind
        acceleration parameter as in the previous cases."

     REVISED
       "Fig. 9 shows the models profiles computed (at $i=55^\circ$)
        using the same ranges of the disc-wind temperature and wind
        acceleration parameter as in the previous cases."

Sect. 4.3, 5th paragraph: It is hard to see in Fig. 10 that the
position of minimum flux in the absorption moves blue-ward. I actually
don not see any difference between the models with mu=0.1 and 0.2.

   * You are correct. I don't know why I said that.
     I rephrase the sentence as follows

     ORIGINAL 
     "The figure shows that as $\mu$ increases the
     P-Cygni absorption deepens only slightly, and the position of minimum
     flux in the absorption trough appears to move blue-ward."

     REVISED
     "The figure shows that as $\mu$ increases the
     P-Cygni absorption deepens only slightly." and the position of minimum
     flux in the absorption trough appears to remain at same locations."

Sect. 5.1, Type III-B: This is the most common profile type in the
Reipurth et al. (1996) sample but you only find it in you Macc = 10^-7
Msun/yr models, which is not a common or average value for Macc rate
in CTTSs. As said in some previous comment, CTTSs tend to have Macc =
10^-8 Msun/yr and only a few present Macc = 10^-7 Msun/yr. To what do
you attribute that discrepancy between the data and the models ?

   * I have run additional models with a typical mass accretion rate
     (3x10^-8 Msun/yr, Gullbring et al. 1998; Calvet et al. 2004) for
     CTTSs.  Please see
     http://www.astro.ex.ac.uk/people/rk/profiles/extra

     for the profile computed at i=10,55,80 and for beta=0.2, 0.5,
     1.0, 2.0 Twind=7000, 8000, 9000, 10000K

      According to these models which use a more typical
      mass-accretion rate, Type III-B can be seen for at least in
      three models with fast accelerating and high wind temperature
      i.e. (beta, Twind) = (0.2, 10000K), (0.5, 9000K) and (0.5,
      10000K).

      A possible reason for not being able to find Type III profiles
      in lower mass-accretion rate models (e.g. in Mdot_acc =
      1.0x10^-8 Msun/yr models) is perhaps due to a wrong radial
      dependence (p in Equation 3) of the mass-loss rate per unit area
      used in our models.  This dependence shall be examined in more
      carefully in future.

      I added a following comments based on these additional runs.

      "In addition to the models shown in Appendix A, we have run a
      set of hybrid models (not shown here) with a typical
      mass-accretion rate for CTTS i.e. 3x10^-8 Msun/yr (Gullbring et
      al. 1998; Calvet et al. 2004) using i=10,55,80 with combinations
      of beta=0.2, 0.5, 1.0, 2.0 and Twind=7000, 8000, 9000,
      10000K. Based on these models, we found that Type III-B profiles
      can be seen for at least in three models with fast accelerating
      and high wind temperature at moderate inclination angle i.e. i =
      55 degrees with (beta, Twind) = (0.2, 10000K), (0.5, 9000K) and
      (0.5, 10000K).  A possible reason for not being able to find
      Type III profiles in lower mass-accretion rate models (e.g. in
      Mdot_acc = 1.0x10^-8 Msun/yr models) is perhaps due to a wrong
      radial dependence (p in Equation 3) of the mass-loss rate per
      unit area used in our models.  This dependence shall be examined
      more carefully in future."

Sect. 5.1, Type IV-R: In axisymmetric models (rotation and magnetic
axis aligned) the inverse P Cygni profile really restricts the
inclination angle to a small range. But there are observational
evidences (SU Aur: Johns & Basri 1995, AA Tau: Bouvier et al. 2003,
for example) for inclined magnetospheres and in that case you could
see the accretion shock through the fast moving accretion flow at
basically any inclination. It would only depend on a combination of
the rotation axis and magnetic axis inclinations. In Halpha, in
particular, it might still be difficult to see inverse P Cygni
profiles because of the wide emission wings due to broadening effects
(such as the Stark effect).

   * Yes, you are quite right. I am aware of the effect of the
     inclined magneto-axis CTTS models. In fact, I wrote a paper on SU
     Aur Pa-beta line variability. H-alpha variability of of SU Aur
     indeed shows a very complex line profile morphology dramatically
     changing with its rotational phase (e.g.  Johns & Basri 1995, see
     also their movie
     http://sprg.ssl.berkeley.edu/~cmj/html/suaur.mpg).  Although it
     does not look it has a clear IV-R type profile at any given
     rotational phase for H-alpha, I am aware that Pa-beta line does
     show IV-R type profile, and the strength of inverse P-Cyg or the
     absorption features in the redwing changes as the rotational
     phase changes (e.g. Kuorsawa et al, 2005, MNRAS, 358, 671).

     I wanted to restrict the cases for purely axi-symmetric
     geometries in this paper for clarity of discussion.  In reality,
     a situation may not be as simple as this classification, as we
     know the lines are sometimes highly variable. This problem will
     be tackled in near future. For a starter, I am currently
     computing line profiles based on the MHD model of Romanova et al
     (2003, ApJ, 595, 1009) who considered the disc accretion problem
     with an inclined dipole magnetic field, and investigate if the
     overall line variability features seen in the observation can be
     explained by this model. I am working on the case with and
     without the additional outflow (discwind or/and stellar wind).

     I will clarify this point by adding the following near the end of
     the first paragraph in Section 5.1 :

       "Although in general H-alpha in the spectra of CTTS shows a
        significant amount of variablity, which is in some cases
        induced by the rotational motion of an inclined magnetosphere
        (e.g. Johns &  Basri 1995; Bouvier et al. 2003), we restrict
        our discussion in this section for the cases of axi-symmetric
        geometries."

    Also a short comment is added at the end of Type IV-R section,

       "The restriction in the inclination angle mentioned above would
        be relaxed if the axis of the magnetosphere is tilted with
        respect to the rotational axis since the line-of-sight to an
        observer from the accretion hot rings will pass through a
        different part of the magnetosphere as a rotational phase
        changes (as it precesses around the rotational axis)."

Sect. 5.2, 3rd paragraph: What is the mass loss rate used in the
stellar wind models ? This value should be added to the text. Matt &
Pudritz (2005) suggest 10^-9 Msun/yr to solve the angular momentum
problem.

   * Thank you for pointing this out.  I should have quoted the
     mass-loss rate used in the models.    
     In both models in Fig.13, Mdot_wind = 10^-8 Msun/yr is used so 
     that mu=Mdot_wind/Mdot_acc = 0.1, and to be consistent with 
     the mass-accretion rate (Mdot_acc=10^-7 Msun/yr) used for the
     models in earlier in the paper. It seems that the ratio mu is an
     important factor in Matt & Pudritz (2005) in balancing the
     torques on a central star (their equation 3).
     I tried the models with Mdot_wind = 10^-9 Msun/yr  (keeping
     mu=0.1), and found the same inclination dependency for these
     models, so the main conclusion of this section 
     does not change when Mdot_wind is changed from 10^-9 Msun/yr to
     10^-9 Msun/yr.

     To reflect your points, I added the following in the text 
     (Sect. 5.2, 4th paragraph).

     After the first sentence of the paragraph, I inserted:

        "The mass-loss rate of the stellar wind in this model is 10^-8
	Msun/yr, which is same as the mass-loss rate used in the
	disc-wind magnetosphere hybrid model in the same figure.  In
	both models, the mass-accretion rate of 10^-7 Msun/yr is
	used."

     At the end of the paragraph, I inserted: 
        "A similar inclination dependency of EW is found for lower
        wind mass-loss rate models (e.g. Mdot_wind = 10^-9 Msun/yr)
        while keeping mu=0.1."

     Also in the figure caption of Figure 13  just before the last
	sentence,  I added:
	"The wind mass-loss rates used in both models are Mdot_wind =
	10^-8 Msun/yr."

***** Spells and other minor comments

Sect. 2.1, 2nd paragraph: ... and re-emitted optical and UV radiation.
Maybe rephrase as: ... and re-emitted as optical and UV radiation.

   * Corrected.

Sect. 2.2, 1st paragraph: Rmi and Rmo --> 2.2 Rsun and 3.0 Rsun should
be 2.2 Rstar and 3.0 Rstar

   * Corrected.

Sect. 2.3, 5th paragraph: The Rs value is set to 10 Rmi. Shouldn't it
be 10 Rwi, as in Table 4 ?

   * Yes. Thank you for noticing. Changed it to Rwi,

Sect. 2.3, caption of Figure 2: ...the density is approximately
proportional to ~ l^2 It looks like ~ l^-2

   * Yes, you are correct. My mistake again. It should be l^-2.
     ==> corrected.

Sect. 3, last paragraph: ...the dust property described in Section 10. 
¨Section 10¨ should be ¨Section 2.4.2¨

   * Corrected.

Sect. 4.2, caption of Figure 7: The 88 Rstar model is represented by a
solid line and not by a dash-dotted line as written.

   * Corrected.

In the same caption, "opening angel" should be "opening angle"

   * Corrected.

Sect. 4.3, 4th paragraph: "Figures 6" should be "Fig. 6"

   * Corrected.

Sect. 4.3, 5th paragraph: ...the flux peak increase and ...  ¨increase¨ 
should be ¨increases¨

   * Corrected.

Sect. 4.3, caption of Figure 11: The parameter used for the
magnetosphere are....  ¨parameter¨ should be ¨parameters¨

   * Corrected.

Sect. 5.2, 4th paragraph: The sentece below must be rewritten: ...This
tendency (...) holds for the models with this wind models for wide
ranges of beta...

   * Changed the sentence 
     From 
     "This tendency of the inclination dependency holds for the models
      with this wind models for wide ranges of beta and
      T_wind providing that the emission from the wind is
      greater than that of the magnetosphere."

     To
     "This inclination dependency of EW is seen in the models with
      wide ranges of beta and T_wind, provided that the wind emission
      of the model is dominating the emission from magnetosphere."

In addition to the referee's suggestions, we noticed the following 
minor mistakes in the original manuscript, and corrected.

1. The mass-loss rate in figure a1 and figure a2 were miss-labeled in
   the captions.  ==>  corrected.

2. In Figure 13, the velocity ranges out the two plots should be
   same for a better comparison. We changed the range of the 
   right plot to -800km/s <V < 800 km/s.
